```{r}
library(STutility)
library(SeuratObject)
library(dplyr)
library(dbscan)
```

## data integration
***

Now we need a strategy to pair data points across the two modalities. 

The strategy used downstream works as follows: For the MSI data pixels, assign at least 1 nearest neighbor 
from the RNA data spots and obtain the average expression for each gene

We also need to apply some kind of distance threshold to remove spots that are too far from a neighbor in 
the other data set.

First, let's calculate the pairwise distances between the two data sets:

```{r include=FALSE}

aligned_coordinates <- readRDS(file = "/Users/jianwu/Documents/gotta/result/aligned_coordinates")

```


```{r eval=FALSE}

# Split coordinates into a list
aligned_coordinates_list <- split(aligned_coordinates, aligned_coordinates$sample_id)

# Find 5 closest spatial neighbors 
knn_spatial_df_filtered_list <- lapply(aligned_coordinates_list, function(coords) {
  
  # get spatial coordinates for RNA and MSI data
  set1 = subset(coords, sample_type == "RNA")[, c("barcode", "warped_x", "warped_y")] |>
    data.frame(row.names = 1)
  set2 = subset(coords, sample_type == "MSI")[, c("barcode", "warped_x", "warped_y")] |>
    data.frame(row.names = 1)
  
  # Calculate pairwise distances using k nearest neighbors algorithm
  knn_spatial_df_filtered <- kNN(x = set1, k = 2, query = set2)
  
  # Create a data.frame with MSI pixel IDs (from) pointing to its closest
  # Visium spot (to) and the distance between the two (distance)
  knn_spatial_df_filtered <- data.frame(from = rownames(set2)[rep(1:nrow(knn_spatial_df_filtered$id), 2)],
                                 to = rownames(set1)[as.vector(knn_spatial_df_filtered$id)],
                                 distance = as.vector(knn_spatial_df_filtered$dist))
  
  # Add spatial coordinates for "from" and "to"
  knn_spatial_df_filtered <- knn_spatial_df_filtered |>
    mutate(x = set2[from, "warped_x"],
           x_end = set1[to, "warped_x"],
           y = set2[from, "warped_y"],
           y_end = set1[to, "warped_y"])
  
  return(knn_spatial_df_filtered)
})
```

What we see in the plot below are the MSI data points in red 
and their closest RNA data neighbors in blue. We need to get 
rid of the MSI data points where we don't have any RNA data points.

```{r fig.width=12, fig.height=30}

plots <- lapply(names(knn_spatial_df_filtered_list), function(nm) {
  knn_spatial_df_filtered <- knn_spatial_df_filtered_list[[nm]]
  p <- ggplot() +
    geom_segment(data = knn_spatial_df_filtered, aes(x = x, xend = x_end, y = y, yend = y_end)) +
    geom_point(data = knn_spatial_df_filtered, aes(x, y), color = "red", size = 0.5) +
    geom_point(data = knn_spatial_df_filtered, aes(x_end, y_end), color = "blue", size = 0.5) +
    scale_y_reverse() +
    labs(title = nm)
  return(p)
})

p <- patchwork::wrap_plots(plots, ncol = 2)
p

```
```{r}
saveRDS(knn_spatial_df_filtered_list,'~/Documents/gotta/result/knn_spatial_df_filtered_list')
```
