
```{r}
se.multi.list <- readRDS('~/Documents/gotta/result/se.multi.list')
```

## Joint analysis
***

NOTE: I selected "CLR" normalization for the MSI data, but another 
normalization method might be more appropriate.

Below, we apply a basic workflow to normalize and analyze the RNA and MSI data sets. The strategy is outlined below:

1. Normalization ("logNormalize" RNA counts, "CLR" for "MSI" values)
2. Find top variable features
3. Run Principal Component Analysis for dimensionality reduction
4. Embed data with UMAP using the PCA results as input (first 30 components for the RNA data and first 10 components for the MSI data)
5. Create an SNN graph from the PCA results (first 30 components for the RNA data and first 10 components for the MSI data)
6. Run data-.driven clustering using default parameters

```{r eval=FALSE}

se.multi.list <- lapply(names(se.multi.list), function(nm) {
  
  se.multi <- se.multi.list[[nm]]
  cat(sprintf("Running analysis for sample %s", nm))
  
  # Normalize RNA data
  DefaultAssay(se.multi) <- 'RNA'
  se.multi <- NormalizeData(se.multi, verbose = FALSE) |> 
    FindVariableFeatures(verbose = FALSE) |> 
    ScaleData(verbose = FALSE) |> 
    RunPCA(reduction.name = "rnapca", verbose = FALSE) |>
    RunUMAP(reduction = "rnapca", dims = 1:30, reduction.name = "RNA.umap", verbose = FALSE) |>
    FindNeighbors(reduction = "rnapca", dims = 1:30, verbose = FALSE) |>
    FindClusters(verbose = FALSE)
  se.multi$RNA_clusters <- se.multi$seurat_clusters
  
  # Normalize MSI data
  DefaultAssay(se.multi) <- 'MSI'
  se.multi <- NormalizeData(se.multi, normalization.method = "CLR", verbose = FALSE) |> 
    FindVariableFeatures(nfeatures = 200, verbose = FALSE) |> 
    ScaleData(verbose = FALSE) |> 
    RunPCA(reduction.name = "msipca", verbose = FALSE) |>
    RunUMAP(reduction = "msipca", dims = 1:10, reduction.name = "MSI.umap", verbose = FALSE) |>
    FindNeighbors(reduction = "msipca", dims = 1:10, verbose = FALSE) |>
    FindClusters(verbose = FALSE)
  se.multi$MSI_clusters <- se.multi$seurat_clusters
  
  return(se.multi)
})

```

We can look at the UMAPs defined by either RNA or MSI data. It is clear that the RNA
data is more complex although it is still possible to find clustered data points in the
MSI data as well.

```{r fig.width=8, fig.height=30}

plots <- lapply(se.multi.list, function(se.multi) {
  p <- DimPlot(se.multi, group.by = "RNA_clusters", reduction = "RNA.umap") | 
    DimPlot(se.multi, group.by = "MSI_clusters", reduction = "MSI.umap")
  return(p)
})

patchwork::wrap_plots(plots, ncol = 1)

```

### Joint clustering
***

see [vignette](https://satijalab.org/seurat/articles/weighted_nearest_neighbor_analysis.html) for more details.

For each cell, we calculate its closest neighbors in the dataset based on a 
weighted combination of RNA and protein similarities. The cell-specific modality 
weights and multimodal neighbors are calculated in a single function, which takes 
~2 minutes to run on this data set. We specify the dimensionality of each modality 
(similar to specifying the number of PCs to include in clustering, but you can 
vary these settings to see that small changes have minimal effect on the overall 
results.

```{r}

# Identify multimodal neighbors. These will be stored in the neighbors slot, 
# and can be accessed using se.multi[['weighted.nn']]
# The WNN graph can be accessed at se.multi[["wknn"]], 
# and the SNN graph used for clustering at se.multi[["wsnn"]]
# Cell-specific modality weights can be accessed at se.multi$RNA.weight

se.multi.list <- lapply(se.multi.list, function(se.multi) {
  se.multi <- FindMultiModalNeighbors(
    se.multi, reduction.list = list("rnapca", "msipca"), 
    dims.list = list(1:30, 1:10), modality.weight.name = "RNA.weight"
  )
  return(se.multi)
})

```

We can now use these results for downstream analysis, such as visualization and 
clustering. For example, we can create a UMAP visualization of the data based on 
a weighted combination of RNA and protein data We can also perform graph-based 
clustering and visualize these results on the UMAP, alongside a set of cell annotations.

```{r eval=FALSE}

se.multi.list <- lapply(se.multi.list, function(se.multi) {
  se.multi <- RunUMAP(se.multi, nn.name = "weighted.nn", reduction.name = "wnn.umap", reduction.key = "wnnUMAP_", n.epochs = 1e3, min.dist = 0.15)
  se.multi <- FindClusters(se.multi, graph.name = "wsnn", algorithm = 3, resolution = 2, verbose = FALSE)
  se.multi$MM_clusters <- se.multi$seurat_clusters
  return(se.multi)
})

```

Now we can also sort the clusters by decreasing RNA weight.

```{r}

plots <- lapply(se.multi.list, function(se.multi) {
  p <- VlnPlot(se.multi, features = "RNA.weight", sort = TRUE)
  return(p)
})

patchwork::wrap_plots(plots[[3]], ncol = 1)

```

And we can compare the results for RNA only, MSI only and RNA+MSI side by side

```{r fig.width=12, fig.height=4}

plots <- lapply(se.multi.list, function(se.multi) {
  p1 <- DimPlot(se.multi, reduction = "RNA.umap", group.by = "RNA_clusters")
  p2 <- DimPlot(se.multi, reduction = "MSI.umap", group.by = "MSI_clusters")
  p3 <- DimPlot(se.multi, reduction = "wnn.umap", group.by = "MM_clusters")
  p <- p1 + p2 + p3 + patchwork::plot_layout(ncol = 3)
  return(p)
})
plots
```

```{r}
saveRDS(se.multi.list,'~/Documents/gotta/result/se.multi.list')
```

